Generality in food webs scales with species richness, not latitude

### 2015-02-03 manuscript/latitudepaper.tex
Up to revising the discussion now

[[re-run with dredge]]
To determine whether there were latitudinal gradients in food-web structure,
we first examined simple linear relationships between latitude and $S$, $Z$, $G$, and $V$. 
We fit models of the form

\begin{eqnarray}
S_{i} = \alpha_{0} + \alpha_{1} L_{i} + \alpha_{2} E_{i} + \alpha_{3} L_{i} E_{i} + \epsilon_{i} 
\end{eqnarray}

where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
latitude (degrees north or south  regardless of direction), $E_{i}$ is the
ecosystem type of network $i$ (comprising terms for stream, marine, lake, and terrestrial networks with estuarine
networks being the intercept) and $\epsilon_{i}$ a residual error term. 
We then simplified these maximal models. To do this, we calculated the AIC
of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models identified
using R~\citep{R} function dredge from packages MuMIn~\citep{MuMIn}. Simplified models were obtained by
systematically removing all possible combinations of terms from the full model.
The best-fitting model was then determined to be the model with the fewest terms where $\delta$AIC<2.

@@

We then assessed
the impact of latitude on this relationship; specifically, the effect of latitude on the scaling exponent $\beta$. 
In addition to latitude, we included a factorial variable for ecosystem type (stream, lake, terrestrial, marine, or estuary) to account for recognized variation in food web structure across ecosystem types \citep{}.
% We also considered other trends in the data that could affect relationships between latitude and species richness.
% To account for changing standards in data collection over time \citep{Dunne2006} we included an effect of year of
% publication. As 11 of our XX webs included humans, and therefore might have different properties than other networks,
% we included a factorial effect of the presence of humans. 
Note that some of our webs describe the same area in different
seasons or different years. 
% [[removed the site-level random effect because I don't want it to steal latitude information.]]

@@

\begin{equation}
\label{PowerLat}
Z_{i}=\alpha S_{i}^{\beta_{0}+\beta_{1}L_{i}+\beta_{3}E_{i}} + \epsilon_{i} 
\end{equation}

where $L_{i}$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
$E_{i}$ indicates the ecosystem type of food web $i$.

@@

\begin{equation}
\label{LogLat}
1log(Z_{i})= log(\alpha)+\beta_{0}log(S_{i}) + \beta_{1}L*log(S_{i}) +\beta_{3}E*log(S_{i}) +\epsilon_{i}
\end{equation}

We fit equation \ref{LogLat}, as well as similar equations for $G$ and $V$ using R \citep{R} function lm from package stats
\citep{stats}.
We calculated the AIC of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models.
Simplified models were obtained by systematically removing all possible combinations of terms from the full model.
The term for species richness was not removed from a model if any of the terms affecting the exponent $\beta$ were 
included in the model. The best-fitting model was then determined to be the model with the fewest terms where $\delta$AIC<2.

The model with the lowest AIC was selected as the best-fitting model. 
This simplification was performed using the
R \citep{R} function dredge from package MuMIn \citep{MuMIn}. We then used the R \citep{R} function lm
to estimate the standardized effects ($\beta$s ) for each fixed effect in the 
best-fitting models, as well as their corresponding $p$-values.

@@

\subsection*{Scaling Relationships with S and Latitude}

For relationships of $Z$, $G$, and $V$ to S, the assumption of multiplicative log-normal error was better supported
(Table \ref{Errorfits}). We therefore use only models with logarithmic form for the remainder of the paper.

Contrary to expected latitudinal gradients, $S$, $Z$, $G$ and $V$ did not vary significantly over latitude, although $S$ and $G$ did vary over ecosystem types (Table~\ref{Latlms}, Fig.~\ref{Latfigs}).

@@

The effect of latitude on scaling relationships varied across ecosystem types
(Table~\ref{Bestfits}, Fig.~\ref{scalingfig}).  In lakes and streams, the
scaling exponent for species richness increased with latitude (although this
increase was not significant for lakes and generality). For other types of
ecosystems, the scaling  exponent for species richness decreased, but not significantly,
with latitude. Despite these overal trends, the observed networks displayed a great deal
of variability in each property.

@@

\begin{center}
\begin{table}[!h]
\caption{Symbols for and descriptions of food-web properties.}
\label{symboltable}
\begin{tabular}{c l l}
\hline
Symbol      & Name & Description \\
\hline
$S$ & Species richness & Number of trophic species \\
$Z$ & Links per species & Mean number of links per taxon \\
$G$ & Generality & Mean number of resources per taxon \\
$V$ & Vulnerability & Mean number of consumers per taxon \\
\hline
\end{tabular}
\end{table}
\end{center}

\begin{table}[!t]

\caption{AIC's for model \ref{Power}, assuming normal error, and model \ref{Loglog}, assuming lognormal error.}
\label{Errorfits}
\begin{tabular}{l | l l | c}
Property & Normal error & Lognormal error & Model selected \\
\hline
$Z$ & 1151 &  767 & \ref{Loglog} \\
$G$ & 1266 & 1037 & \ref{Loglog} \\
$V$ & 1151 & 767 & \ref{Loglog} \\
\end{tabular}
\end{table}

\begin{table}[!h]
\caption{Intercepts and slopes for simple linear regressions of food web properties against absolute latitude.}
\label{Latlms}
\begin{tabular}{l | l l | l l}

\multirow{2}{*}{Property} & \multicolumn{2}{|c}{Intercept} & \multicolumn{2}{|c}{Slope} \\
                      & $\alpha_{0}$ & $p$-value & $\alpha_{1}$ & $p$-value \\
\hline
$S$ & 29.42 & \textless0.001 & 0.188 & 0.167 \\
$Z$ & 2.724 & 0.002          & 0.017 & 0.283 \\
$G$ & 3.095 & 0.006          & 0.035 & 0.080 \\
$V$ & 2.724 & 0.002          & 0.017 & 0.283 \\
\end{tabular}
\end{table}

\begin{table}[!h]

\caption{Standard effects and their $p$-values for the best-fitting models of links per species ($Z$), generality ($G$), and vulnerability ($V$).}
\label{Bestfits}
\begin{tabular}{l | l l | l l | l l }
Effect          & \multicolumn{2}{c|}{$Z$} & \multicolumn{2}{c|}{$G$} & \multicolumn{2}{c}{$V$}\\
                & $\beta$ & $p$-value     &  $\beta$ & $p$-value    &  $\beta$ & $p$-value  \\
\hline
Intercept               & -0.561 & \textless0.001 & -0.299 & \textless0.001 & -0.561 & \textless0.001 \\
log(species)            &  0.733 & \textless0.001 &  0.622 & \textless0.001 &  0.733 & \textless0.001 \\
log(species):Lake       & -0.118 & 0.047          & -0.115 & 0.078          & -0.118 & 0.047 \\
log(species):Marine     &  0.051 & 0.040          &  0.120 & \textless0.001 &  0.051 & 0.040 \\
log(species):Stream     & -0.161 & 0.014          & -0.057 & 0.421          & -0.161 & 0.014 \\
log(species):Latitude   & -0.001 & 0.081          & -0.001 & 0.273          & -0.001 & 0.081 \\
log(species):Lat:Lake   &  0.006 & \textless0.001 &  0.006 & \textless0.001 &  0.006 & \textless0.001 \\
log(species):Lat:Stream &  0.004 & 0.018          &  0.003 & 0.059          &  0.004 & 0.018 \\
\hline
$R^{2} (adjusted)$   & \multicolumn{2}{c|}{0.706} & \multicolumn{2}{c|}{0.658} & \multicolumn{2}{c}{0.706} \\
\end{tabular}
\end{table}

@@

\begin{figure}[!h]
\label{Latfigs}
\includegraphics[width=.9\textwidth]{Figures/properties_vs_lat.eps}
\caption{Relationships between latitude and species richness, mean links per species, mean generality, and mean vulnerability. No food web property showed a significant latitudinal gradient in this dataset (N=259 food webs).}
\end{figure}

\begin{figure}[!h]
\label{scalingfig}
\includegraphics[width=.9\textwidth]{Figures/scaling_with_S_grey.eps}
\caption{Relationships between species richness and mean links per species, mean generality, and mean vulnerability (N=259 food webs). In estuarine, terrestrial, and marine communities the scaling relationship with species-richness did not vary significantly across latitude for any food-web property. In lake and stream communities, each food-web property showed stronger scaling with species richness as latitude increased (N=259 food webs). }
\end{figure}

>>>>

To determine whether there were latitudinal gradients in food-web structure,
we first examined simple linear relationships between latitude and $S$, $Z$, $G$, and $V$. 
We fit models of the form

\begin{equation}
\label{Latfull}
S_{i} = \alpha_{0} + \alpha_{1} L_{i} + \alpha_{2} E_{i} + \alpha_{3} L_{i} E_{i} + \epsilon_{i} 
\end{equation}

where $S_{i}$ is the species richness of web $i$, $L_{i}$ its absolute
latitude (degrees north or south  regardless of direction), $E_{i}$ is the
ecosystem type of network $i$ (comprising terms for stream, marine, lake, and terrestrial networks with estuarine
networks being the intercept) and $\epsilon_{i}$ a residual error term. 
We then simplified these maximal models. To do this, we calculated the AIC
of the chosen maximal model, as well as the AIC's of a suite of candidate simplified models identified
using R~\citep{R} function dredge from packages MuMIn~\citep{MuMIn}. Simplified models were obtained by
systematically removing all possible combinations of terms from the full model.
The best-fitting model was then determined to be the model with the fewest terms 
where $\Delta$AIC\textless2. If several models shared the fewest number of terms 
and had $\Delta$AIC\textless2, the model with the lowest AIC in that set was chosen as the best-fit
model.

@@

We then assessed the impact of latitude on the scaling relationships between $Z$, $G$, $V$, and $S$. 
Even if there is no direct effect of latitude on specialization, perhaps due to contrasting effects
on species richness as described above, there may still be an effect of latitude on the scaling of
each property with species richness. In fact, if the species-rich tropics truly have narrower 
niches~\citep{Brown2004}, then we would expect to see less increase in mean numbers of links, 
predators, and prey with increasing species richness in the tropics.
In the context of the scaling relationships above, we wish to determine the effect of latitude on
the scaling exponent $\beta$. As when examining the relationships between latitude and each food 
web property directly, we included a factorial variable for ecosystem type (stream, lake, 
terrestrial, marine, or estuary), as well as interactions between food web type and latitude.

@@

\begin{equation}
\label{PowerLat}
Z_{i}=\alpha S_{i}^{\beta_{0}+\beta_{1}L_{i}+\beta_{2}E_{i}+\beta_{3}LE_{i}} + \epsilon_{i} 
\end{equation}

where $S_{i}$ is the number of species in food web $i$, $L_{i}$ is the absolute latitude of food web $i$ (degrees from the equator, regardless of direction),
$E_{i}$ indicates the ecosystem type of food web $i$.

@@

\begin{equation}
\label{LogLat}
\log{Z_{i}} = \log{\alpha}+\beta_{0}\log{S_{i}} + \beta_{1}L\log{S_{i}} +\beta_{2}E\log{S_{i}} +\beta_{3}LE\log{S_{i}} +\epsilon_{i}
\end{equation}

For each property in $Z$, $G$, and $V$, we used the form of the equation that was best supported
when describing the scaling of the property with species richness alone~\citep{Xiao2011}.
We then simplified the model. As with the models describing the relationships between latitude and food-web
properties, we calculated the AIC of the maximal and all reduced models, created by systematically removing
all possible combinations of terms. Note that if an interaction term was included in a reduced model, all
associated main effects were also retained. $S$ was included in all reduced models. Reduced models were
identified using the R~\citep{R} function dredge from package MuMIn~\citep{MuMIn}. The best-fitting model
was then determined to be the model with the fewest terms where $\Delta$AIC<2. Where there were several
models with the fewest terms where $\Delta$AIC<2, the model with the lowest AIC in this set was chosen as
the best-fitting model. The best-fitting model was then fit using R~\citep{R} function lm from package stats~\citep{stats}.

@@

\subsection*{Scaling Relationships with $S$ and Latitude}

For relationships of $Z$, $G$, and $V$ to $S$, equation~\ref{LogLog} had a lower AIC than the corresponding
equation~\ref{Power}. This indicates that the data support an assumption of multiplicative log-normal
error better than an assumption of additive normal error. That is, [[but what does it mean?]] . 
We therefore used logarithmic-form models when assessing the effect of latitude on scaling relationships 
with $S$.

Contrary to the expected latitudinal gradient, the best-fit version of equation~\ref{Latfull} for $S$
did not include a term for latitude or any of its interactions with ecosystem type. Species richness was
higher for stream ecosystems ($\alpha_{2:Stream}$=22.4, $p$\textless0.001) than other ecosystem types. 
Similarly, the best-fit version of equation~\ref{Latfull} for $G$ did not include 
any terms for latitude or interactions between latitude and ecosystem type. $G$ was lower in terrestrial
and lake food webs than in other ecosystem types ($\alpha_{2:Terrestrial}$=-3.77, $p$\textless0.001 and 
$\alpha_{2:Lake}$=-1.98, $p$=0.044). The best-fit models for $Z$ and $V$ did not include any terms for
latitude, ecosystem type, or their interactions.
the best-fit version of equation~\ref{Latfull} for $Z$ did not include a term for latitude,
ecosystem type, or any of their interactions. Food webs tended to contain about 3.53 links per species
($p$\textless0.001), with the average species being preyed upon by 3.53 predators ($p$\textless0.001).

@@

The effect of latitude on scaling relationships varied across ecosystem types.
Overall, specialization in marine and terrestrial food webs increased towards the poles, 
while specialization in lake food webs increased towards the equator (Fig.~\ref{scalinfig}). 
In estuarine and stream food webs, there was a great deal of variation, with links per species
and vulnerability increasing towards the poles and generality increasing towards the equator.
More specifically, the scaling of mean links per species with species richness increased with
increasing latitude ($\beta_{Latitude}$=0.004, $p$\textless0.001; Table~\ref{Bestfits}). This
increase, however, was reversed for marine and terrestrial food webs
($\beta_{Latitude:Marine}$=-0.005, $p$=0.007 and
$\beta_{Latitude:Terrestrial}$=-0.005, $p$=0.001). Scaling of mean vulnerability with
species richness showed the same pattern ($\beta_{Latitude}$=0.004, $p$\textless0.001;
$\beta_{Latitude:Marine}$=-0.005, $p$=0.007; and $\beta_{Latitude:Terrestrial}$=-0.005, $p$=0.002).
In contrast, scaling of mean generality with species richnesss increased with latitude 
for lake food webs ($\beta_{Latitude:Lake}$=0.006, $p$=0.002) but did not vary with latitude for
other ecosystem types ($\beta_{Latitude}$=-0.001, $p$=0.199).

@@

\begin{table}[!h]
\caption{Intercepts and slopes for the best-fitting model describing the relationship between species-richness, mean links per species, mean generality, and mean vulnerability and latitude. Note that none
of the best-fitting models included a term for latitude or an interaction between latitude and ecosystem type. Each `NA' indicates a term that was not included in the corresponding best-fit model.}
\label{Latlms}
\begin{tabular}{l | l l  l l  l l  l l}

\hline
\multirow{2}{*}{Property} & \multicolumn{2}{|c}{Intercept} & \multicolumn{2}{|c}{Stream} & \multicolumn{2}{|c}{Lake} & \multicolumn{2}{|c}{Terrestrial}\\
& $\alpha_{0}$ & $p$-value & $\alpha_{2}$ & $p$-value & $\alpha_{2}$ & $p$-value & $\alpha_{2}$ &$p$-value\\
\hline
Species richness  & 34.6 & \textless0.001 & 22.4 & 0.167 \textless0.001 & NA & NA & NA & NA \\
Links per species & 3.53 & \textless0.001 & NA & NA & NA & NA & NA & NA \\
Generality        & 7.32 & \textless0.001 & NA & NA & -1.98 & 0.044 & -3.77 & \textless0.001 \\
Vulnerability     & 3.53 & \textless0.001 & NA & NA & NA & NA & NA & NA \\
\hline
\end{tabular}
\end{table}

\begin{table}[!h]
\caption{Intercepts and slopes for the best-fitting model describing the relationship between species-richness, mean links per species, mean generality, and mean vulnerability and latitude in trophic-species versions of food webs. Note that none
of the best-fitting models included a term for latitude or an interaction between latitude and ecosystem type. `NA' a term that was not included in the corresponding best-fit model.}
\label{Latlms}
\begin{tabular}{l | l l  l l  l l  l l}

\hline
\multirow{2}{*}{Property} & \multicolumn{2}{|c}{Intercept} & \multicolumn{2}{|c}{Stream} & \multicolumn{2}{|c}{Lake} & \multicolumn{2}{|c}{Terrestrial}\\
& $\alpha_{0}$ & $p$-value & $\alpha_{2}$ & $p$-value & $\alpha_{2}$ & $p$-value & $\alpha_{2}$ &$p$-value\\
\hline
Species richness  & 34.6 & \textless0.001 & 22.4 & 0.167 \textless0.001 & NA & NA & NA & NA \\
Links per species & 3.53 & \textless0.001 & NA & NA & NA & NA & NA & NA \\
Generality        & 7.32 & \textless0.001 & NA & NA & -1.98 & 0.044 & -3.77 & \textless0.001 \\
Vulnerability     & 3.53 & \textless0.001 & NA & NA & NA & NA & NA & NA \\
\hline
\end{tabular}
\end{table}

\begin{table}[!h]
\caption{Intercepts and slopes for best-fit models describing the effect of latitude on the scaling of mean links per species, mean generality, and mean vulnerability with species richness. Each `NA' indicates a term that was not included in the corresponding best-fit model. These models refer to the original versions of food webs. }
\label{Bestfits}
\begin{tabular}{l | l l  l l  l l }
Effect          & \multicolumn{2}{c|}{Links per species} & \multicolumn{2}{c|}{Generality} 
                & \multicolumn{2}{c}{Vulnerability}\\
                & $\beta$ & $p$-value     &  $\beta$ & $p$-value    &  $\beta$ & $p$-value  \\
\hline
Intercept               &-0.453 & \textless0.001 & -0.249 & \textless0.001 & -0.453 & \textless0.001 \\
log(species)            & 0.419 & \textless0.001 &  0.663 & \textless0.001 &  0.419 & \textless0.001 \\
log(species):Lake       & 0.119 & \textless0.001 & -0.185 & 0.016          &  0.119 & \textless0.001 \\
log(species):Marine     & 0.255 & \textless0.001 &  0.090 & 0.002          &  0.255 & \textless0.001 \\
log(species):Terrestrial& 0.207 & 0.002          & -0.122 & \textless0.001 &  0.207 & 0.002 \\
log(species):Latitude   & 0.004 & \textless0.001 & -0.001 & 0.199          &  0.003 & \textless0.001 \\
log(species):Latitude:Marine&-0.004 & 0.007      &  NA    &  NA            & -0.005 & 0.007 \\
log(species):Latitude:Terrestrial&-0.005 & 0.002 &  NA    &  NA            & -0.005 & 0.002 \\
log(species):Latitude:Lake & NA & NA             &  0.006 & 0.002          &  NA    & NA    \\
\hline
$R^{2} (adjusted)$   & \multicolumn{2}{c|}{0.706} & \multicolumn{2}{c|}{0.658} & \multicolumn{2}{c}{0.706} \\
\hline
\end{tabular}
\end{table}

\begin{table}[!h]
\caption{Intercepts and slopes for best-fit models describing the effect of latitude on the scaling of mean links per species, mean generality, and mean vulnerability with species richness. Each `NA' indicates a term that was not included in the corresponding best-fit model. These models refer to the trophic species versions of food webs. }
\label{Bestfits}
\begin{tabular}{l | l l  l l  l l }
Effect          & \multicolumn{2}{c|}{Links per species} & \multicolumn{2}{c|}{Generality} 
                & \multicolumn{2}{c}{Vulnerability}\\
                & $\beta$ & $p$-value     &  $\beta$ & $p$-value    &  $\beta$ & $p$-value  \\
\hline
Intercept               & -0.561 & \textless0.001 & -0.299 & \textless0.001 & -0.561 & \textless0.001 \\
log(species)            &  0.733 & \textless0.001 &  0.622 & \textless0.001 &  0.733 & \textless0.001 \\
log(species):Lake       & -0.118 & 0.047          & -0.115 & 0.078          & -0.118 & 0.047 \\
log(species):Marine     &  0.051 & 0.040          &  0.120 & \textless0.001 &  0.051 & 0.040 \\
log(species):Stream     & -0.161 & 0.014          & -0.057 & 0.421          & -0.161 & 0.014 \\
log(species):Latitude   & -0.001 & 0.081          & -0.001 & 0.273          & -0.001 & 0.081 \\
log(species):Lat:Lake   &  0.006 & \textless0.001 &  0.006 & \textless0.001 &  0.006 & \textless0.001 \\
log(species):Lat:Stream &  0.004 & 0.018          &  0.003 & 0.059          &  0.004 & 0.018 \\
\hline
$R^{2} (adjusted)$   & \multicolumn{2}{c|}{0.706} & \multicolumn{2}{c|}{0.658} & \multicolumn{2}{c}{0.706} \\
\hline
\end{tabular}
\end{table}

@@

% \begin{figure}[!h]
% \label{Latfigs}
% \includegraphics[width=.9\textwidth]{Figures/properties_vs_lat.eps}
% \caption{Relationships between latitude and species richness, mean links per species, mean generality, and mean vulnerability. No food web property showed a significant latitudinal gradient in this dataset (N=259 food webs).}
% \end{figure}

\begin{figure}[!h]
\label{scalingfig}
\includegraphics[width=.9\textwidth]{Figures/scaling_with_S_grey.eps}
\caption{Relationships between species richness and mean links per species, mean generality, and mean 
vulnerability (N=259 food webs). Scaling of mean links per species and mean vulnerability with species 
richness strengthened with latitude for estuarine, stream, and lake food webs. This strengthening was
reversed for marine and terrestrial food webs. Scaling of mean generality with species richness 
strengthened with latitude for lake food webs, but did not significantly vary with latitude for other
ecosystem types (N=259 food webs). }
\end{figure}
